---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)

library(gapminder)
data("gapminder")
attach(gapminder)

gapminder <- read.csv("data/gapminder_data.csv", header = TRUE)
head(gapminder)

```

```{}
gap <-gapminder
rm(gapminder)

mean(gap[gap$continent == "Africa", "gdpPercap"])


mean.gdpAfrica <- mean(gap$gdpPercap[continent=="Africa"])
mean.gdpAfrica
mean.gdpAsia <- mean(gap$gdpPercap[continent=="Asia"])
mean.gdpAsia
mean.gdpAmericas <- mean(gap$gdpPercap[continent=="Americas"])
mean.gdpAmericas
```

```{}
ncol(gap)
names(gap)
subset_gap <- %>% select (year, country, gdpPercap)
```

```{}
year_country_gdp_euro <- gap %>%
fliter(continent =="Europe") %>%
select(year, country, gdpPercap)

year_country_gdp_euro

str(year_country_gdp_euro)
```

```{}
year_country_lifeExp_Africa <- gap %>%
filter(continent == "Africa", year == 2007) %>%
select(year, country, lifeExp)

dim(year_country_lifeExp_Africa)
```

```{}
head(gap)
str(gap)
gap_by_continent <- gap %>% group_by(continent)
head(gap_by_continent)
str(gap_by_continent)

```
group_by() turns a single data frame into a set of data frames grouped based on some variable.
```{}
gdp_by_continent <- gap_by_continent %>%
summarize(mean_gdpPercap=mean(gdpPercap))

gdp_by_continent 

```

```{}
Africa_lifeExp_yr <- gap %>%
filter(continent == "Africa") %>%
group_by(year)  %>%
summarize(avg_life = mean(lifeExp))
options(pillar.sigfig = 4)
Africa_lifeExp_yr

```

calculate the average life expectancy per country. Which has the longest avearge life expectancy> and which has the shoretst average life expectancy?

```{}
life <- gap %>%
group_by(country) %>%
summarize(mean_lifeExp = mean(lifeExp))

life %>% filter(mean_lifeExp == min(mean_lifeExp))

life %>% filter(mean_lifeExp == max(mean_lifeExp))

```
mutate()is a function in dplyr that lets you add a column to a data frame based the data
```{}
billgdp <- gap %>%
filter (year==2007)%>%
mutate(billion_gdp = gdpPercap * pop / 10^9) %>%
select(continent, country, billion_gdp)

colnames(billgdp)
dim(billgdp)
```

```{}
gdp_by_cont <- gap %>%
group_by(continent,year) %>%
summarize(mean_gdp = mean(gdpPercap),
sd_gdp=sd(gdpPercap),
mean_pop=mean(pop),
sample_size=n(),
se_gdp=sd_gdp/sqrt(sample_size))

gdp_by_cont

str(gdp_by_cont)

gap_by_cont <- gdp_by_cont %>% data.frame()


str(gdp_by_cont)

```
#function n()counts sample size of the var grouped by

```{}
library(ggplot2)
gap %>% filter(continent == "Asia") %>%
ggplot(aes(x= gdpPercap, y=lifeExp))+
geom_point()
```

```{}
gap_wide <- read.csv("data/gapminder_wide.csv",stringsAsFactors = FALSE)
options(stringsAsFactors = FALSE)

dim(gap_wide)
colnames(gap_wide)
```

```{}
head(gap_wide)
library(tidyr)
## To make the data long, we use the function gather()
gap_long <- gap_wide %>%
gather(obstype_year, obs_values, -continent, -country)
dim(gap_long)
```

```{}
gap_long <- gap_long %>% separate(obstype_year, into = c("obs_type", "year"), sep="_")
head(gap_long)

gap_long$year <- as.integer(gap_long$year)
head(gap_long)


```
Using the gap_long dataset, summarize the mean life expectancy by continent.

```{}
gap_long %>% filter(obs_type == "lifeExp")%>%
group_by(continent) %>%
summarize(mean_lifeExp = mean(obs_values))

temp_gap <- filter(gap_long, obs_type == "lifeExp")
temp_gap <- group_by(temp_gap, continent)
temp_gap <- summarize(temp_gap, mean_lifeExp = mean(obs_values))
temp_gap 

str(temp_gap)

```

```{}
str(gap_long)
gap_normal <- gap_long %>% spread(obs_type, obs_values)
str(gap_normal)
```

colnames(gap_normal)
str(gap_normal)

str(gap)
colnames(gap)

```{}
names(gap)
gap_normal <- gap_normal [, names(gap)]

names(gap)
names(gap_normal)

all.equal(gap, gap_normal)

head(gap)
head(gap_normal)
all.equal(gap, gap_normal)
str(gap)
str(gap_normal)
```
